---
title: "data-story-reflection"
date: today
format: 
  html:
    self-contained: true
    toc: true
    toc_float: true
    number_section: false
    highlight: "tango"
    theme: "cosmo"
editor: visual
editor_options: 
  chunk_output_type: console
---

# Reflection

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(olsrr)
cleaned_screen_time_data = readRDS("./project/data/cleaned_screen_time_data.rds")
```

> The story that emerged from our preliminary analysis is that screen time negatively impacts mental wellness and productivity by increasing stress, with students being the most vulnerable group. Looking at the correlation matrix, screen time shows a moderate negative relationship with both mental wellness and productivity while showing a positive relationship with stress. This gave us a clear narrative thread to pull on: screen time feeds stress, and stress is what's quietly eroding people's productivity and wellbeing. The regression model put a concrete number on it, with every additional hour of screen time associated with roughly a 3.7-point drop in mental wellness, which felt like the kind of finding that could anchor the whole presentation.

> That said, we identified a few gaps worth addressing before treating this as a complete story. The most significant is causality: our data shows correlations, not proof that screen time drives these outcomes, and a skeptical audience could reasonably argue the relationship runs the other way. The closing point around sleep also might need to be a stronger bridge connecting it back to screen time specifically, otherwise it risks feeling like a topic switch rather than a resolution. 

> Our next steps are to subset the data to students and refit the key models to see if the relationships hold, explore whether screen time before bed might explain the sleep quality connection more directly, and consider pulling out just the screen time row from the correlation matrix into a cleaner standalone visual that's easier for a general audience to read at a glance.

# Visualizations

### Plot 1: Who is on their screens the most?

```{r}
cleaned_screen_time_data |>
  summarise(avg_screen_time = mean(screen_time_hours), .by = occupation) |>
  ggplot(aes(x = avg_screen_time, y = reorder(occupation, avg_screen_time))) +
  geom_col() +
  labs(
    x = "Average Screen Time (Hours)",
    y = "Occupation",
    title = "Average Screen Time by Occupation"
  )
```

> Students spend more time on screens than any other group in this dataset, which makes them the natural focus for the rest of this story. But spending more time on screens raises an obvious question: what is all that screen time actually doing to them?

---

### Plot 2: The Bigger Picture

```{r}
st_data_without_categories = cleaned_screen_time_data[,c(2, 6, 9:15)]
st_data_without_categories |> 
  cor() |> 
  ggcorrplot(lab = TRUE)
```

> The correlation matrix shows that screen time is moderately negatively correlated with mental wellness and productivity, and positively correlated with stress. So it is not just one outcome being affected — screen time seems to touch several areas of wellbeing at once. Stress stands out as a likely connector, so let's look at that more closely.

---

### Plot 3: Productivity vs Stress

```{r}
cleaned_screen_time_data |> 
  ggplot(aes(x = productivity_0_100, y = stress_level_0_10)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Productivity",
    y = "Stress Level",
    title = "Stress in Relation to Productivity"
  )
```

> Higher stress clearly corresponds with lower productivity. This helps connect the dots: if screen time is pushing stress up, then it is also indirectly pulling productivity down. Now that we can see how stress fits into the picture, let's put a real number on how much damage screen time is doing overall.

---

### Plot 4: Linear Regression Model

```{r}
wellness_model <- lm(mental_wellness_index_0_100 ~ exercise_minutes_per_week + 
                       screen_time_hours + sleep_quality_1_5,
                     data = cleaned_screen_time_data)

summary(wellness_model)
```

> The model estimates that every additional hour of screen time is associated with roughly a 3.7-point drop in mental wellness, even after accounting for exercise and sleep quality. That is a meaningful hit. But the model also tells us something hopeful: sleep quality is one of the strongest predictors of wellness in the model, which gives us somewhere to look for solutions.

---

### Plot 5: Sleep as a Solution

```{r}
cleaned_screen_time_data |> 
  ggplot(aes(x = sleep_hours, y = sleep_quality_1_5)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Hours of Sleep",
    y = "Quality of Sleep",
    title = "Sleep Quality in Relation to Hours of Sleep"
  )
```

> People who sleep more tend to report better sleep quality, and better sleep quality is one of the strongest predictors of mental wellness in our model. Rather than ending on a warning, this gives us something actionable: one of the most effective things someone can do to protect their wellbeing may simply be to prioritize getting enough sleep, especially if cutting screen time feels out of reach.